Towards cross-middleware authentication and single sign-on for ARC Grid middleware

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
More minor correction

%\authorrunning{Short form of author list} % if too long for running head
\institute{W. Qiang \at
              Department of Physics \\
              University of Oslo, Norway\\
              \email{weizhong.qiang@fys.uio.no}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           A. Konstantinov \at
              \email{aleksandr.konstantinov@fys.uio.no}
}
\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor
\maketitle
\begin{abstract}
When pursuing the task of making access to Grids as simple as possible, security is one of the most important challenges in production Grid infrastructures, especially when the Grid applications span multiple administrative domains as well as heterogeneous Grid middlewares. A typical example is wide scale e-Science applications which need to coordinate resources shared among a number of independent institutions with different Grid middlewares deployed on these resources. In this paper, we describe security implementation and considerations used in the upcoming version of the Advanced Resource Connector (ARC) middleware, where the heterogeneity issue has been addressed. The main goal of ARC implementation in terms of security is to let the middleware be capable of interoperating with other Grid middlewares by leveraging on standard specifications. The key aspect of the work is to enhance the current proxy certificate based authentication and single sign-on by utilizing and enhancing the standardized Web Service specifications such as Security Assertion Markup Languages (SAML), single sign-on (SSO) profile and Web Services Security in order to achieve cross-middleware authentication and single sign-on.
\keywords{single sign-on \and delegation \and virtual organization \and ARC middleware }
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)~\cite{AlfieriR05} have partly solved them, to a different extent. The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.  As the most successful Grid authorization model, VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach. The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity. Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.

@@

Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
\begin{itemize}
\item The GSS API based confidential communication has not been adopted by Web Service implementations.
\item The Attribute Certificate has not been adopted by any aspects Web Service implementation at all.
\item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
\item Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another. But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.
\end{itemize}

@@

The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users’ daily experience and middleware’s requirements in order to ensure the access as simple as possible.

@@

The rest of this paper is organized as follows: Section II presents the ARC Grid middleware, especially the architecture of the new version of ARC. Section III describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on. Section IV discusses the related research, and Section V presents conclusions and future work outlook.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED.png}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler -- one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the "AuthZ" and "AuthN" sub-modules inside MCCs and Services are examples of SecHandler.

@@

The Shibboleth provides cross-domain single sign-on and attribute-based authorization while preserving user privacy. It is based on the OASIS Security Assertion Markup Language (SAML), specifically, the new version of Shibboleth supports SAML 2.0 specification. For the authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines two functional components, an Identity Provider and a Service Provider. The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request. In order to discover which home community does a user come from, Shibboleth specifies an optional third component called ``Where Are You From?'' (WAYF) service to aid in the process of IdP discovery. This IdP discovery process is also standardized and defined in SAML 2.0 specification as “Identity Provider Discovery Profile”.

@@

\section{Conclusion And Future Work}
\label{sec:conclusion}
In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middleware authentication and single sign-on. The goal of our work is to develop an interoperable and accessible single sign-on solution to be used for a VO scenario in production Grids, where users can easily authenticate against the Grids with any of their credentials (not only X.509) from anywhere, and move inside Grid middleware infrastructure and even across heterogeneous Grids  boundaries, without re-authentication within a reasonable period. In order to do so, we implemented some standards based Web Services (including the SLCS service, the X.509 credential delegation service, and the SAML Token delegation service), as well as the user agent and the Service Provider in the context of the SAML2 SSO profile.

>>>>

%\authorrunning{Short form of author list} % if too long for running head
\institute{W. Qiang \at
              Department of Physics \\
              University of Oslo, Norway\\
              \email{weizhong.qiang@fys.uio.no}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           A. Konstantinov \at
              Department of Physics \\
              University of Oslo, Norway\\
              \email{aleksandr.konstantinov@fys.uio.no}
}
\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor
\maketitle
\begin{abstract}
When pursuing the task of making access to Grids as simple as possible, security is one of the most important challenges in production Grid infrastructures, especially when the Grid applications span multiple administrative domains as well as heterogeneous Grid middlewares. A typical example is wide scale e-Science applications which need to coordinate resources shared among a number of independent institutions with different Grid middlewares deployed on these resources. In this paper, we describe security implementation and considerations used in the upcoming version of the Advanced Resource Connector (ARC) middleware, where the heterogeneity issue has been addressed. The main goal of ARC implementation in terms of security is to let the middleware be capable of interoperating with other Grid middlewares by leveraging on standard specifications. The key aspect of the work is to enhance the current proxy certificate based authentication and single sign-on by utilizing and enhancing the standardized Web Service specifications such as Security Assertion Markup Languages (SAML), single sign-on (SSO) profile and Web Services Security in order to achieve cross-middleware authentication and single sign-on.
\keywords{single sign-on \and delegation \and virtual organization \and ARC middleware }
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS) ~\cite{AlfieriR05} have partly solved them, to a different extent. The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.  As the most successful Grid authorization model, VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach. The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity. Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.

@@

Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
\begin{itemize}
\item The GSS API based confidential communication has not been adopted by Web Service implementations.
\item The Attribute Certificate has not been adopted by any aspects Web Service implementation at all.
\item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
\item Finally, identity delegation is a good solution for user or user's representative (acting on behalf of the user) to move/travel from one resource to another. But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.
\end{itemize}

@@

The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users' daily experience and middleware's requirements in order to ensure the access as simple as possible.

@@

The rest of this paper is organized as follows: Section 2 presents the ARC Grid middleware, especially the architecture of the new version of ARC. Section 3 describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on. Section 4 discusses the related research, and Section 5 presents conclusions and future work outlook.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED.png}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler -- one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside MCCs and Services are examples of SecHandler.

@@

The Shibboleth provides cross-domain single sign-on and attribute-based authorization while preserving user privacy. It is based on the OASIS Security Assertion Markup Language (SAML), specifically, the new version of Shibboleth supports SAML 2.0 specification. For the authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines two functional components, an Identity Provider and a Service Provider. The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request. In order to discover which home community does a user come from, Shibboleth specifies an optional third component called ``Where Are You From?'' (WAYF) service to aid in the process of IdP discovery. This IdP discovery process is also standardized and defined in SAML 2.0 specification as ``Identity Provider Discovery Profile''.

@@

\section{Conclusion And Future Work}
\label{sec:conclusion}
In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middle-ware authentication and single sign-on. The goal of our work is to develop an interoperable and accessible single sign-on solution to be used for a VO scenario in production Grids, where users can easily authenticate against the Grids with any of their credentials (not only X.509) from anywhere, and move inside Grid middleware infrastructure and even across heterogeneous Grids  boundaries, without re-authentication within a reasonable period. In order to do so, we implemented some standards based Web Services (including the SLCS service, the X.509 credential delegation service, and the SAML Token delegation service), as well as the user agent and the Service Provider in the context of the SAML2 SSO profile.
